---
title: "Assignment 3"
author: "Luis Sanchez Mercedes"
date: "9/15/2021"
output: html_document
---


Problems
-------
 \

1. Create the following data frame

| Rank 	| Age 	| Name  	|
|------	|-----	|-------	|
| 0    	| 28  	| Tom   	|
| 1    	| 34  	| Jack  	|
| 2    	| 29  	| Steve 	|
| 3    	| 42  	| Ricky 	|

```{r}
df <- data.frame(Rank = c(0, 1, 2, 3),
                 Age  = c(28, 34, 29, 42),
                 Name = c('Tom', 'Jack', 'Steve', 'Ricky'))
df
```

2. Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv)

Show the names of the variables in the data
```{r}
df_covid <- read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')
names(df_covid)

```


3. How many columns and rows the data have? 
```{r}
str(df_covid)
```
print('')

4. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?
```{r}
# Display the total amount of missing
sum(is.na(df_covid))

```

```{r}
# Display the toal amount of missing by column (Note: all columns have zero missing)
colSums(is.na(df_covid))
```


5. What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column.
```{r}
# Display the type of the date variable (Note: it is a string)
class(df_covid$date)
```
```{r}
# Convert the type of the date variable from string into date and display type again
df_covid$date = as.Date(df_covid$date)
class(df_covid$date)
```
6. Capitalize the names of all the variables
```{r}
names(df_covid)[1] <- "Date"
names(df_covid)[2] <- "State"
names(df_covid)[3] <- "Fips"
names(df_covid)[4] <- "Cases"
names(df_covid)[5] <- "Deaths"

# Display after modification at hand
str(df_covid)
```

7. Find the average number of cases per day. Find the maximum cases a day.  
```{r}
# Find the mean number of cases
mean(df_covid$Cases, na.rm = TRUE)
```
```{r}
# Find the maximum number of cases
max(df_covid$Cases, na.rm = TRUE)
```

8. How many states are there in the data?  
```{r}
# Display a table with all the states included (Note: 50 States + 5 protected territories )
table(df_covid$State)
```


9. Create a new variable `weekdays` to store the weekday for each rows. 
```{r}
df_covid$weekdays <- weekdays(df_covid$Date)
table(df_covid$weekdays)
```



10. Create the categorical variable `death2` variable taking the values as follows

- `has_death` if there is a death that day
-  `no_death` if there is no death that day
```{r}
df_covid$death2 <- ifelse(df_covid$Deaths==0, 'no_death', 'has_death')
```


Find the frequency and relative frequency of `no_death` and `has_death`. 
```{r}
table(df_covid$death2)
```

```{r}
prop.table(table(df_covid$death2))
```


11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint:  Use the `summary` function)
```{r}
# Q1 is 363, Q2 (the median) is 2079, Q3 is 7371
summary(df_covid$Deaths)

```


12. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3
```{r}
df_covid$death3 <- cut(df_covid$Deaths,  
                       breaks=c(-Inf, 363, 7371, Inf), 
                       labels=c('low_death', 'mid_death', 'high_death'))
table(df_covid$death3)
```

13. Find the average cases in Rhode Island in 2021
```{r}
df_covid_ri21 <- df_covid[(df_covid$State=='Rhode Island')&(df_covid$Date>='2021-01-01'),]
summary(df_covid_ri21$Deaths)[4]
```

14. Find the median cases by weekdays in Rhode Island in 2021 
```{r}
by(df_covid_ri21$Deaths, df_covid_ri21$weekdays, median)
```

15. Compare the median cases in Rhode Island in June, July, August and September. 
```{r}
# Note: Most median cases in September
df_covid_ri21$month <- months(df_covid_ri21$Date)
df_covid_ri21 <- df_covid_ri21[(df_covid_ri21$Date>='2021-06-01'),]
by(df_covid_ri21$Deaths, df_covid_ri21$month, median)

```


