---
title: "Assignment 4"
author: "Luis Sanchez Mercedes"
date: "9/29/2021"
output: html_document
---

```{r}
library(tidyverse)
```


#### 2. Read the data using `read_csv`

Use `read_csv` function to import the US Covid 19 data at [link](https://covidtracking.com/data/download/all-states-history.csv).  Don't forget to import `tidyverse` (library(tidyverse)) so that you can use read_csv. 
```{r}
# DATA-001: Read csv file from site
df <- read_csv('https://covidtracking.com/data/download/all-states-history.csv')
```


-------

#### 3.  Fix the date and ceate some new variables

`lubridate` is a package of the `tidyverse` packages.  We will make uses of `lubridate` in this question.

- Use the following codes to change the type of the `date` column to `date`. 

```{r}
library(lubridate)
df$date <- ymd(df$date)
```

- Use the below codes to create `month`, `weekday` and `monthday` variables

```{r}

df$month = month(df$date)

# day of the week
df$weekday = wday(df$date)

# day of the month
df$monthday <- mday(df$date)
```

-------

#### 4. Create new variables with `case_when`. 

The function `case_when` is a good option to create a new variable from existing variable.  For example, this below codes create a new variable, `daily_death`, from `deathIncrease` variable.  `deathIncrease` is the number of daily new death by Covid19.  The new variable `daily_death` takes three values:  low (if `deathIncrease` less than 3), medium (`deathIncrease` from 3 to 14), and high (`deathIncrease` more than 14).  Please notice that this can also be done in a different way as shown in Assignment 3. 

- Create variable `month2` that takes three values: early_month (day of the month from 1-10), mid_month (day of the month from 11-20), and end_month (day of the month > 20).
```{r}
df$month2 <- case_when(
  df$monthday <=10 ~ 'early_month',
  df$monthday <=20 ~ 'mid_month',
  TRUE ~ 'end_month'
)

table(df$month2)
```


- Create variable `weekend` that takes two values:  1 if it's Saturday or Sunday or 0 otherwise. 
```{r}
df$weekend <- case_when(
  df$weekday <= 5 ~ 0,
  TRUE ~ 1
)
```
```{r}
table(df$weekend)
```


-------

#### 5. Select function

Use the select function to deselect the column `totalTestsViral` from the data.
```{r}
df <- select(df, -totalTestsViral)
```

```{r}
df
```



-------

#### 6. Pipe Operator ( %>% )

Pipe operator offers another way to write R codes. Many times, it makes the codes more readable.  Pipe works very well with all the `tidyverse` packages. Refer to [these slides](file:///C:/Users/sonou/Dropbox/git/math421/slides/dplyr.html#15) (slide 15, 16, 17 and 18) to rewrite the below codes using pipe operator

```{r, eval=FALSE}
x <- c(1:10) 

# square root of x

x %>% sum() 

x %>% sum() %>%  sqrt()

x %>% sum() %>%  sqrt() %>% log()

# log base 2 of 16
log(16, 2)

```

-------

#### 7. Combo 1: group_by + summarise

This combo is used when you want to apply a function/calculation to different groups of the data. For example, to calculate the average number of cases (`positiveIncrease`) by `dataQualityGrade`, we use:

```{r, eval=FALSE}
df %>% 
  group_by(weekday) %>% 
  summarise(mean(positiveIncrease))
```


- Calculate the median number of cases (`positiveIncrease`) by `month`
```{r}
df %>% 
  group_by(month) %>% 
  summarise(mean(positiveIncrease))
```


- Calculate the average number of cases (`positiveIncrease`) by `month2`
```{r}
df %>% 
  group_by(month2) %>% 
  summarise(mean(positiveIncrease))
```


- Calculate the median number of cases (`positiveIncrease`) by `weekend`
```{r}
df %>% 
  group_by(weekend) %>% 
  summarise(mean(positiveIncrease))
```



-------

#### 8. Combo 2: filter + group_by + summarise

An example: to calculate the average number of cases (`positiveIncrease`) in January and February separately, we use:

```{r, eval=FALSE}
df %>% 
  filter(month==1|month==2) %>% 
  group_by(month) %>% 
  summarise(positve_increase = mean(positiveIncrease))
```

- Calculate the median number of cases (`positiveIncrease`) on the weekend by `month` in October and November 2020. 
```{r}
df$year <- year(df$date)
table(df$year)
```




```{r}
df %>% 
  filter(month==10|month==11) %>% 
  filter(year==2020) %>% 
  group_by(month) %>% 
  summarise(positve_increase = mean(positiveIncrease))
```


- Calculate the average number of death at different periods of a month (`month2` variable) in Fall 2020
```{r}
df %>% 
  filter(month>=9) %>% 
  filter(year==2020) %>% 
  group_by(month2) %>% 
  summarise(death_increase = mean(deathIncrease))
```


- Compare the average number of hospitalizations between weekdays and weekends in Summer 2020
```{r}
df %>% 
  filter(month>4 & month<8) %>% 
  filter(year==2020) %>% 
  group_by(weekend) %>% 
  summarise(hospitalization_increase = mean(hospitalizedIncrease))
```

- Redo Questions 14 and 15 in [Assignment 3](fa2021_assignment3.html) using the combos.  Notice:  you also need to use the data used in Assignment 3.  
```{r}
df_covid <- read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')
df_covid$date = as.Date(df_covid$date)
df_covid$year <- year(df_covid$date)
df_covid$month <- month(df_covid$date)
df_covid$weekday <- weekdays(df_covid$date)

```



```{r}
# Question 14
df_covid %>% 
  filter(year==2021) %>% 
  filter(state=='Rhode Island') %>% 
  group_by(weekday) %>% 
  summarise(avg_case = mean(cases))
```

```{r}
# Question 15 (No Results because 2021 only had January, February and March)
df_covid %>% 
  filter(year==2021) %>%
  filter(month>5, month<10) %>%
  filter(state=='Rhode Island') %>% 
  group_by(month) %>% 
  summarise(median_case = mean(cases))
```


-------

#### 9. Combo 2: filter + group_by + summarise + arrange

Use the arrange function to find a month that has the highest number of deaths on the weekend.
```{r}
df %>% 
  filter(weekend==1) %>% 
  group_by(month) %>% 
  summarise(sum_death = sum(death)) %>% 
  arrange(-sum_death) %>% 
  head(n=1)

```

