---
title: "Assignment 9"
author: "Luis Sanchez Mercedes"
date: "10/20/2021"
output: html_document
---

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 


```{r}
library(gganimate)
library(ggplot2)
library(tidyverse)
library(lubridate)

df_covid <- read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')
df_covid
```


2. Make a plot revealing the number of deaths in the three countries with the most numbers of deaths
```{r}

top_country <- df_covid %>% 
                filter(Date_reported==Sys.Date()) %>% 
                  group_by(Country) %>% 
                    summarise(sum = sum(Cumulative_deaths)) %>% 
                     mutate(rank=rank(-sum)) %>% 
                     filter(rank<=3) %>% 
                pull(Country)

df_covid %>% 
  filter(Country %in% top_country) %>% 
    ggplot(aes(y=Cumulative_deaths, x=Date_reported, color=Country))+
    geom_line()+
    transition_reveal(Date_reported)
  
```


3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 
```{r}
df_covid$death_per_cases <- df_covid$Cumulative_deaths/df_covid$Cumulative_cases


top_country_dpc <- df_covid %>% 
                     group_by(Country) %>% 
                       summarise(mean = mean(death_per_cases)) %>% 
                        mutate(rank=rank(-mean)) %>% 
                        filter(rank<=3) %>% 
                     pull(Country)

top_country_dpc
```


4. Make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 
```{r}
df_covid %>% 
  filter(Country %in% c('United States of America', 'Italy', 'Mexico')) %>% 
    ggplot(aes(y=death_per_cases, x=Date_reported, color=Country))+
    geom_line()+
    transition_reveal(Date_reported)

```


5. Use `transition_reveal` to create three animated graph with the vaccination data at 

https://raw.githubusercontent.com/govex/COVID-19/master/data_tables/vaccine_data/us_data/time_series/vaccine_data_us_timeline.csv
```{r}
df_vaccine<- read_csv('https://raw.githubusercontent.com/govex/COVID-19/master/data_tables/vaccine_data/us_data/time_series/vaccine_data_us_timeline.csv')
df_vaccine

```
```{r}
df_vaccine$month <-  month(df_vaccine$Date)


df_vaccine %>% 
  filter(Vaccine_Type %in% c('Moderna', 'Pfizer')) %>%
   group_by(month, Vaccine_Type) %>% 
    summarize(sum=sum(Doses_admin, na.rm =TRUE)) %>% 
      ggplot(aes(y=sum, x=month, color=Vaccine_Type))+
        geom_line()+
        transition_reveal(month)

```


```{r}
new_england <- c('Maine', 'Vermont', 'New Hampshire', 'Massachusetts', 'Connecticut', ' Rhode Island')


df_vaccine %>% 
  filter(month>5, month<10) %>% 
  filter(Province_State %in% new_england) %>%
   group_by(month, Province_State) %>% 
    summarize(sum=sum(Doses_admin)) %>% 
      ggplot(aes(y=sum, x=month, color=Province_State))+
        geom_line()+
        transition_reveal(month)
```

```{r}
northeast = c('Connecticut', 'Maine', 'Massachusetts', 'New Hampshire', 'Rhode Island', 'Vermont', 'New Jersey', 'New York', 'Pennsylvania')
midwest   = c('Illinois', 'Indiana', 'Michigan', 'Ohio', 'Wisconsin', 'Iowa', 'Kansas', 'Minnesota', 'Missouri', 'Nebraska', 'North Dakota', 'South Dakota')
west      = c('Arizona', 'Colorado', 'Idaho', 'Montana', 'Nevada', 'New Mexico', 'Utah', 'Wyoming', 'Alaska', 'California', 'Hawaii', 'Oregon', 'Washington')

df_vaccine$US_Region <- 'South'
df_vaccine$US_Region[df_vaccine$Province_State %in% northeast] <- 'Northeast'
df_vaccine$US_Region[df_vaccine$Province_State %in% midwest]   <- 'Midwest'
df_vaccine$US_Region[df_vaccine$Province_State %in% west]      <- 'West'


table(df_vaccine$US_Region)
```

```{r}
df_vaccine %>% 
  filter(month>5, month<10) %>% 
   group_by(month, US_Region) %>% 
    summarize(sum=sum(Doses_admin)) %>% 
      ggplot(aes(y=sum, x=month, color=US_Region))+
        geom_line()+
        transition_reveal(month)
```

