---
title: "Assignment 11"
author: "Luis Sanchez Mercedes"
date: "10/27/2021"
output: html_document
---


1. Install the package `mlbench` and use the follows to import the data

```{r}
library(mlbench)
library(caret)


data(PimaIndiansDiabetes)
df_diabetes <- PimaIndiansDiabetes
```

- Set seed to be 2020. 
```{r}
set.seed(2020)
```

- The target variable is `diabetes`
```{r}
df_diabetes$target <- df_diabetes$diabetes
df_diabetes$diabetes <- NULL
```

- Partition the data into 80% training and 20% testing.  
```{r}
splitIndex <- createDataPartition(df_diabetes$target, p=.80, list=FALSE )

df_train <- df_diabetes[splitIndex, ]
df_test  <- df_diabetes[-splitIndex,]
df_train
```


-------

2. Tuning Decision Tree:  Use cross-validation with 10 k-folds to find the maxdepth with the greatest accuracy. Plot the accuracy associated with different maxdepths against the maxdepths. The range to search for maxdepth is from 1 to 10. 
```{r}
tuning  = expand.grid(maxdepth = 1:10)
control = trainControl(method = "cv", number = 10)

decision_tree <- train(target~., 
                        data=df_train, 
                          method = "rpart2", 
                           trControl = control,
                           tuneGrid = tuning)
```


```{r}
plot(decision_tree)
```

-------

3. Make the final decision to select the maxdepth for your decision tree.  Is your selected maxdepth the same as the maxdepth found in 2? 
```{r}
decision_tree2 <- train(target~., data=df_train, method = "rpart2", maxdepth=1)
```


-------

4. Calculate the accuracy of your decision tree (the decision tree with your selected maxdepth in 3) on the test data. 
```{r}
pred <- predict(decision_tree2, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = "neg")
cm$overall[1]
```

